---
title: "Q5"
output: html_document
date: "2023-12-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*Import the data for double-stranded DNA (dsDNA) viruses taken from the Supplementary Materials of the original paper into Posit Cloud (the csv file is in the question-5-data folder). How many rows and columns does the table have? (3 points)*  

```{r}
row_number <- nrow(Cui_etal2014)
row_number #33
column_number <- ncol(Cui_etal2014)
column_number #13
```
*What transformation can you use to fit a linear model to the data? Apply the transformation. (3 points)*  
You can log transform the data by applying the natural logarithm to the data.
```{r}
# Applying the log transformation:
columns_to_log <- c("Virion volume (nm×nm×nm)", "Genome length (kb)") #We only want to transform Virion volume and Genome length, so we remove the other columns
transformation <- log(Cui_etal2014[, columns_to_log])
```

*Find the exponent (α) and scaling factor (β) of the allometric law for dsDNA viruses and write the p-values from the model you obtained, are they statistically significant? Compare the values you found to those shown in Table 2 of the paper, did you find the same values? (10 points)*  
```{r}
linear_model <- lm(transformation) #Fitting linear model to data
exponent <- coef(linear_model)[2]
scaling_factor <- exp(coef(linear_model)[1])

cat("Exponent (alpha):", exponent, "\n")
cat("Scaling Factor (beta):", scaling_factor, "\n")

summary(linear_model) #obtaining p-values
```
*Write the code to reproduce the figure shown below. (10 points)*  
```{r}
# Install and load packages
install.packages("ggplot2")
library(ggplot2)

# Code to reproduce the figure
ggplot(linear_model, aes(x = `Virion volume (nm×nm×nm)`, y = `Genome length (kb)`)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, fill = "grey58", colour = "royalblue2", linewidth = 0.7) + #adding a blue linear regression line and a grey 95% cl
  labs(x = "log [Genome length (kb)]", y = "log [Virion volume nm3)]") +
  theme_minimal() + 
  theme(axis.title = element_text(face = "bold")) + #adding bold font to axis titles
  theme(panel.border = element_rect(colour = "grey42", fill = NA, size = 0.5))
```

*What is the estimated volume of a 300 kb dsDNA virus? (4 points)*
```{r}
log300 <- log(300) #finding log of 300 to align with the log transformed model 
log300
estimate_volume <- (1.5152*log300) + (7.0748) #using equation of a straight line (y = mx+c) to solve for estimated volume where m = 1.5152 (slope estimate), x = log300 (genome length) and c = 7.0748 (intercept estimate)
estimate_volume #answer = 15.71717

```

